---
title: "hw 3"
author: "Lulu Zhang"
date: "10/8/2019"
output: github_document
---


--
__Problem 1__
--


```{r}
library(tidyr)
library(tidyverse)
library(dplyr)
library(ggplot2)
library(httr)
library(jsonlite)
# load the data
library(p8105.datasets)
data("instacart") 

```

```{r}
#How many aisles are there, and which aisles are the most items ordered from?
instacart
instacart %>% 
  count(aisle) %>% 
  arrange(desc(n))
  
# there are 134 aisles
# fresh vegetables and fresh fruits are two of the most ordered from aisles

```

  
Make a plot that shows the number of items ordered in each aisle, limiting this to aisles 
with more than 10000 items ordered. Arrange aisles sensibly, and organize your plot so others can read it.
```{r}
plot1 = instacart %>%
   count(aisle) %>% 
  arrange(desc(n)) %>% 
  filter(n > 10000) %>% 
           ggplot(aes(x = aisle, y = n)) +
           geom_point(aes(color = n)) #%>% +
 plot1 + theme(axis.text.x = element_text(angle = 90,  hjust = 1)) 
  
```

Make a table showing the three most popular items in each of the aisles “baking ingredients”, 
“dog food care”, and “packaged vegetables fruits”. Include the number of times each item is ordered in your table.
```{r}
#make a plot 
instacart %>% 
  count(aisle)
 # order(product_name, descending = FALSE)
```

  
#Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream 
#are ordered on each day of the week; format this table for human readers (i.e. produce a 2 x 7 table).

```{r}
#make a table

table = instacart %>% 
  janitor::clean_names() %>% 
  sort(product_name, by == "Pink Lady Apples") %>% 
  mean(order_hour_of_day)
  pivot_longer(product_name,
               names_to = "product_name",
               values_to = "order_hour_of_day")
  table
```

--
__Problem 2__
--

First, do some data cleaning:

format the data to use appropriate variable names;
focus on the “Overall Health” topic
include only responses from “Excellent” to “Poor”
organize responses as a factor taking levels ordered from “Poor” to “Excellent”
Using this dataset, do or answer the following (commenting on the results of each):

```{r}
library(p8105.datasets)
data("brfss_smart2010")

brfss = brfss_smart2010 %>% 
  janitor::clean_names() %>% 
  filter(topic == "Overall Health") %>% 
  drop_na(response) %>% 
  mutate(
    response = as.factor(response),
   # colnames(brfss)[colnames(brfss) =="locationabbr"] <- "state"
    #recode(state = locationabbr)
  ) %>% 
  select(-data_value_footnote, -location_id, -data_value_footnote_symbol) 

brfss
```

In 2002, which states were observed at 7 or more locations? What about in 2010?

```{r}
brfss %>% 
  filter(year == "2002") %>% 
  n_distinct(locationdesc, na.rm = FALSE)
```


Construct a dataset that is limited to Excellent responses, and contains, year, state, and a variable that averages the data_value across locations within a state. Make a “spaghetti” plot of this average value over time within a state (that is, make a plot showing a line for each state across years – the geom_line geometry and group aesthetic will help).
Make a two-panel plot showing, for the years 2006, and 2010, distribution of data_value for responses (“Poor” to “Excellent”) among locations in NY State.

```{r}
excellent = brfss %>% 
  filter(response == "Excellent") %>% 
  select(year, )
```





